---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# phantSEM

<!-- badges: start -->
[![R-CMD-check](https://github.com/argeorgeson/phantSEM/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/argeorgeson/phantSEM/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of phantSEM is to make it easier to create phantom variables, which are variables that were not observed, for the purpose of sensitivity analyses for structural equation models. The package allows a user to test different combinations of covariances between the phantom variable(s) and observed variables. 


## Installation
You can install phantSEM from CRAN:
```{r, eval=FALSE}
install.packages("phantSEM")
```


You can install the development version of phantSEM from [GitHub](https://github.com/argeorgeson/phantSEM) with:

``` r
# install.packages("devtools")
devtools::install_github("argeorgeson/phantSEM")
```


## Example

This is an example that shows you how to use the package.  Assume that you have a simple mediation model with three variables, X, M2, and Y2. You can use the phantSEM package to create phantom variables for baseline observations of M and Y. 

```{r example}
library(phantSEM)
## basic example code
covmatrix <- matrix(c(
  0.25, 0.95, 0.43,
  0.95, 8.87, 2.66,
  0.43, 2.66, 10.86
), nrow = 3, byrow = TRUE)
colnames(covmatrix) <- c("X", "M2", "Y2")

# lavann syntax for observed model
observed <- " M2 ~ X
              Y2 ~ M2+X "

# lavaan output
obs_output <- lavaan::sem(model = observed, sample.cov = covmatrix, sample.nobs = 200)

summary(obs_output)
# lavaan syntax for phantom variable model
phantom <- " M2 ~ M1 + Y1 + a*X
                Y2 ~ M1 + Y1 + b*M2 + cp*X "

Step1 <- SA_step1(
  lavoutput = obs_output,
  mod_obs = observed,
  mod_phant = phantom
)

phantom_assignment <- list(
  "CovM1X" = 0,
  "CovY1M1" = "CovY2M2",
  "CovY1X" = 0,
  "VarM1" = 1,
  "VarY1" = 1,
  "CovM1M2" = seq(0, .6, .1),
  "CovY1Y2" = "CovM1M2",
  "CovY1M2" = seq(-.6, .6, .1),
  "CovM1Y2" = "CovY1M2"
)
Step2 <- SA_step2(
  phantom_assignment = phantom_assignment,
  step1 = Step1
)
Step3 <- SA_step3(
  step2 = Step2,
  n = 200
)

b_results <- ghost_par_ests(
  step3 = Step3,
  parameter_label = "b",
  remove_NA = TRUE
)

head(b_results)
```
## Help
If you encounter errors in the package, please submit an issue with a minimal reproducible example on Github or email the package maintainer. 

